Automated Verification and Synthesis of Embedded Systems: Advances and Challenges

### 2017-08-02 paper.tex
TCPS format

\documentclass{acm_sen_article}

@@

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{Lucas C. Cordeiro \\
%Department of Computer Science\\
University of Oxford, UK\\
lucas.cordeiro@cs.ox.ac.uk
\and
Eddie B. de Lima Filho \\
Samsung Electronics, Brazil \\
eddie.l@samsung.com
\and
Iury V. de Bessa \\
Federal University of Amazonas, Brazil \\
iury.bessa@ufam.edu.br}
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678--2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

@@

%\IEEEpeerreviewmaketitle

@@

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems}. Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).

@@

%--------------------------------------
\section{A model for Cyber-physical systems}
\label{sec:model}
%--------------------------------------

@@

CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata (\textcolor{red}{CITAR}) and hybrid automata (\textcolor{red}{CITAR}). Rungger and Tabuada (\textcolor{red}{CITAR TAC 2016}) proposed a general and wide model for CPS. Following their proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.

\textcolor{red}{seria legal falar que nao determinismo pode ser tratado com model checking e ligar com o nosso trabalho... inclusive , o tabuada tem um artigo falando sobre seguranca e SMT, que eu comentei abaixo}

@@

@@

Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.

@@

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific program classes, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal to remove redundant behaviour or analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.

@@

@@

%1
\bibitem{Kopetz11}
H.~Kopetz:
\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.

%2
\bibitem{xu15}
Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.

%3
\bibitem{leeCPS}
Lee E.:
\newblock {\em {Cyber-physical Systems: Design Challenges}}.
\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.

%4
\bibitem{leeCPS2}
Lee E.:
\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
\newblock In: Sensors 15(3): pp. 4837--4869, 2015.

%5
\bibitem{groza2015formal}
Groza A., Letia I., Goron A., Zaporojan S.: 
\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.

%6
\bibitem{Cordeiro09}
Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.

%7
\bibitem{munir}
Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.

%8
\bibitem{Abate17}
Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).

%9
\bibitem{ldd}
Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
\newblock {\em {Model checking concurrent linux device drivers}}.
\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.

%10
\bibitem{mlembe}
S.~Virtanen:
\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.

%11
\bibitem{Kroening15}
Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.

%12
\bibitem{cudalucas}
Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.

%13
\bibitem{stochmodel}
Bargmann H.:
\newblock {\em {The role of stochastic modelling in engineering science}}.
\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.

%14
\bibitem{Shoukry}
Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.

%15
\bibitem{cpsattack}
Liu Y., Ning P., Reiter M. K.:
\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.

%16
\bibitem{Araujo16}
Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).

%17
\bibitem{Biere99}
Biere A., Cimatti A., Clarke E., Zhu Y.:
\newblock {\em {Symbolic Model Checking without BDDs}}. 
\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.

%18
\bibitem{handbook09}
Biere A., Heule M., van Maaren H., Walsh T., eds.:
\newblock {\em {Handbook of Satisfiability}}.
\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.

%19
\bibitem{BarrettSST09}
Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
\newblock {\em {Satisfiability Modulo Theories}}. 
\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.

%20
\bibitem{Armando06}
Armando A., Mantovani J., Platania L.:
\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.

%21
\bibitem{Bradley07}
Bradley A., Manna Z.:
\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.

%22
\bibitem{Thomas01}
Thomas Ball and Sriram~K. Rajamani.
\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.

%23
\bibitem{Strichman08}
Daniel Kroening and Ofer Strichman.
\newblock {\em Decision Procedures: An Algorithmic Point of View}.
\newblock Springer Publishing Company, Incorporated, 2008.

%24
\bibitem{Appel97}
Andrew~W. Appel.
\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
\newblock Cambridge University Press, New York, NY, USA, 1997.

%25
\bibitem{Tseitin83}
G.~S. Tseitin.
\newblock On the complexity of derivation in propositional calculus.
\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.

 %26
 \bibitem{PatarinG97}
Jacques Patarin and Louis Goubin.
\newblock Trapdoor one-way permutations and multivariate polynominals.
\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997.

%27
\bibitem{Clarke04}
Clarke E., Kroening D., Lerda F.:
\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.

%28
\bibitem{MerzFS12}
Merz F., Falke S., Sinz C.:
\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.

%29
\bibitem{CordeiroF11}
Cordeiro L., Fischer B.:
\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.

%30
\bibitem{Ivancic05}
Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.

%31
\bibitem{Cordeiro12}
Cordeiro L., Fischer B., Marques{-}Silva J.:
\newblock {\em {{SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}}.
\newblock {IEEE} Trans. Software Eng. 38(4), pp. 957--974, 2012.

>>>>

\documentclass[format=acmsmall, review=false, screen=true]{acmart}

\usepackage{booktabs} % For formal tables
% \usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{hyperref}	
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{xcolor,colortbl}
% \usepackage{cite}
\newtheorem{mydef}{Definition}
\usepackage{multirow}
\usepackage{xspace}
\usepackage[ruled]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmJournal{TCPS}
\acmVolume{9}
\acmNumber{9}
\acmArticle{99}
\acmYear{2017}
\acmMonth{12}
\copyrightyear{2017}
%\acmArticleSeq{9}

% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{0000001.0000001}

% Paper history
\received{February 2007}
\received[revised]{March 2009}
\received[accepted]{June 2009}

@@

\author{Lucas C. Cordeiro}
\orcid{0000-0002-6235-4272}
\affiliation{%
  \institution{Department of Computer Science, University of Oxford}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Oxford}
%  \state{VA}
%  \postcode{23185}
  \country{UK}}

\author{Eddie B. de Lima Filho}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{TP Vision}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

\author{Iury V. de Bessa}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{Department of Electricity, Federal University of Amazonas}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

%% author names and affiliations
%% use a multiple column layout for up to three different
%% affiliations
%\author{Eddie B. de Lima Filho \\
%%Department of Computer Science\\
%University of Oxford, UK\\
%lucas.cordeiro@cs.ox.ac.uk
%\and
%Eddie B. de Lima Filho \\
%Samsung Electronics, Brazil \\
%eddie.l@samsung.com
%\and
%Iury V. de Bessa \\
%Federal University of Amazonas, Brazil \\
%iury.bessa@ufam.edu.br}
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678--2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

@@

%\IEEEpeerreviewmaketitle
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010405.10010432.10010439.10010440</concept_id>
<concept_desc>Applied computing~Computer-aided design</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010553</concept_id>
<concept_desc>Computer systems organization~Embedded and cyber-physical systems</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010178.10010213.10010214</concept_id>
<concept_desc>Computing methodologies~Computational control theory</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10002944.10011122.10002945</concept_id>
<concept_desc>General and reference~Surveys and overviews</concept_desc>
<concept_significance>100</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Computer-aided design}
\ccsdesc[500]{Computer systems organization~Embedded and cyber-physical systems}
\ccsdesc[300]{Computing methodologies~Computational control theory}
\ccsdesc[100]{General and reference~Surveys and overviews}

@@

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs). Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).

@@

%--------------------------------------
\section{Automated Verification and Synthesis of Cyber-physical systems}
\label{sec:model}
%--------------------------------------

Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS. A CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

The overall process of formal and automated synthesis for CPS and HS involves three main steps: modeling, that is related to obtain a sound mathematical representation that is able to capture all the relevant aspects of continuous and discrete dynamics and their interaction, as well as formulate properties and invariants that describe the desirable behavior; verification, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller synthesis itself, that produces a controller or finds invariant control laws that are able to make the CPS, modeled according to the first step (modeling), meets the requirements and behaves as desired, as confirmed by the verification step.

%--------------------------------------
\subsection{A model for Cyber-physical systems}
\label{ssec:model}
%--------------------------------------

@@

CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.

Timed automata employ finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories with constant slope. Hybrid automata generalize timed automata, also employing finite automata with real variables whose trajectories follow more general dynamical laws~\cite{Henzinger95}.  Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.

Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for CPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for CPS.

%Following their proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,t)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.

%\textcolor{red}{seria legal falar que nao determinismo pode ser tratado com model checking e ligar com o nosso trabalho... inclusive , o tabuada tem um artigo falando sobre seguranca e SMT, que eu comentei abaixo}

%--------------------------------------
\subsection{Automated and Formal Verification of Cyber-physical systems}
\label{ssec:verification}
%--------------------------------------

Given a class of CPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in CPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model CPS) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must be hold for CPS are related to safety, liveness, and reachability. In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.} bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

@@

%--------------------------------------
\subsection{Control synthesis of Cyber-physical systems}
\label{ssec:synthesis}
%--------------------------------------

@@

Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}.

@@

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific program classes, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal to remove redundant behaviour or analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

@@

%1
%\bibitem{Kopetz11}
%H.~Kopetz:
%\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
%\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
%
%%2
%\bibitem{xu15}
%Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
%\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
%\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
%
%%3
%\bibitem{leeCPS}
%Lee E.:
%\newblock {\em {Cyber-physical Systems: Design Challenges}}.
%\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
%
%%4
%\bibitem{leeCPS2}
%Lee E.:
%\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
%\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
%
%%5
%\bibitem{groza2015formal}
%Groza A., Letia I., Goron A., Zaporojan S.: 
%\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
%\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
%
%%6
%\bibitem{Cordeiro09}
%Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
%\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
%\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
%
%%7
%\bibitem{munir}
%Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
%\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
%\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
%
%%8
%\bibitem{Abate17}
%Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
%\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
%\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
%
%%9
%\bibitem{ldd}
%Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
%\newblock {\em {Model checking concurrent linux device drivers}}.
%\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
%
%%10
%\bibitem{mlembe}
%S.~Virtanen:
%\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
%\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.
%
%%11
%\bibitem{Kroening15}
%Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
%\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
%\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
%
%%12
%\bibitem{cudalucas}
%Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
%\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
%\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.
%
%%13
%\bibitem{stochmodel}
%Bargmann H.:
%\newblock {\em {The role of stochastic modelling in engineering science}}.
%\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.
%
%%14
%\bibitem{Shoukry}
%Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
%\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
%\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.
%
%%15
%\bibitem{cpsattack}
%Liu Y., Ning P., Reiter M. K.:
%\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
%\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.
%
%%16
%\bibitem{Araujo16}
%Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
%\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
%\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
%
%%17
%\bibitem{Biere99}
%Biere A., Cimatti A., Clarke E., Zhu Y.:
%\newblock {\em {Symbolic Model Checking without BDDs}}. 
%\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
%
%%18
%\bibitem{handbook09}
%Biere A., Heule M., van Maaren H., Walsh T., eds.:
%\newblock {\em {Handbook of Satisfiability}}.
%\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
%
%%19
%\bibitem{BarrettSST09}
%Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
%\newblock {\em {Satisfiability Modulo Theories}}. 
%\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
%
%%20
%\bibitem{Armando2009}
%Armando A., Mantovani J., Platania L.:
%\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
%\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
%
%%21
%\bibitem{Bradley07}
%Bradley A., Manna Z.:
%\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
%\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
%
%%22
%\bibitem{Thomas01}
%Thomas Ball and Sriram~K. Rajamani.
%\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
%\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
%
%%23
%\bibitem{Strichman08}
%Daniel Kroening and Ofer Strichman.
%\newblock {\em Decision Procedures: An Algorithmic Point of View}.
%\newblock Springer Publishing Company, Incorporated, 2008.
%
%%24
%\bibitem{Appel97}
%Andrew~W. Appel.
%\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
%\newblock Cambridge University Press, New York, NY, USA, 1997.
%
%%25
%\bibitem{Tseitin83}
%G.~S. Tseitin.
%\newblock On the complexity of derivation in propositional calculus.
%\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
%  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
% 
% %26
% \bibitem{PatarinG97}
%Jacques Patarin and Louis Goubin.
%\newblock Trapdoor one-way permutations and multivariate polynominals.
%\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997. 
%
%%27
%\bibitem{Clarke04}
%Clarke E., Kroening D., Lerda F.:
%\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.
%
%%28
%\bibitem{MerzFS12}
%Merz F., Falke S., Sinz C.:
%\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
%\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
%
%%29
%\bibitem{CordeiroF11}
%Cordeiro L., Fischer B.:
%\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
%\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
%
%%30
%\bibitem{Ivancic05}
%Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
%\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
%\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
%
%%31
